---
title: "Simpsons_Emery_cooking"
author: "Emery Lauer"
date: "2025-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#install.packages("plotly")
```

## EDA

```{r}
df_ep <- read.csv("data/simpsons_episodes.csv")
df_script <- read.csv("data/simpsons_script_lines.csv")

head(df_ep)
head(df_script)
```

#### Compute the total number of words per episode & plot relative to viewership

```{r}
# Clean up df_ep to make it slightly easier to join down the line (ie, have the join by colum be named the same thing)
# Also filter out extra columns that are not needed
 df_ep_new <- df_ep %>% 
  select(id, imdb_votes, original_air_date, original_air_year, us_viewers_in_millions, views) %>% 
  rename(
    episode_id = id
  )

# Compute the total number of works in an episode
df_words_per_epsiode <-
  df_script %>% 
  filter(speaking_line == TRUE) %>% 
  mutate(word_count = as.numeric(word_count)) %>%
  filter(word_count <= 1000) %>% 
  group_by(episode_id) %>% 
  summarise(total_words_per_episode = sum(word_count)) %>% 
  inner_join(df_ep_new, by = "episode_id")

df_words_per_epsiode %>%
  ggplot(aes(total_words_per_episode, us_viewers_in_millions)) +
  geom_point()
```

#### Compute words over time

```{r}
df_words_per_epsiode %>%
  ggplot(aes(episode_id, total_words_per_episode)) +
  geom_point() +
  geom_smooth()
```

#### Compute Speech Complexity

```{r}
unique_words <- function(line)
  # Convert all characters to lowercase
  tolower(line) %>% 
  # Remove all non-alphanumeric characters (ie punctuation)
  gsub("[^[:alnum:] ]", "", .) %>%
  # Split into individual words
  str_split(" +") %>%
  # Deal with weird list-in-a-list thing 
  .[[1]] %>% 
  # Count onlthe unique (non-repeat words)
  unique() %>% 
  # Get total number of words
  length()

df_speech_complexity <-
  df_script %>%
  mutate(word_count = as.numeric(word_count)) %>%
  filter(word_count < 1000) %>%
  filter(
    grepl(
      "Simpson",
      raw_character_text
    )
  ) %>%
  group_by(raw_character_text) %>%
  summarise(
    all_words = paste0(normalized_text, collapse = " "),
    num_unique_words = unique_words(all_words),
    total_words = sum(word_count, na.rm = TRUE),
    speech_complexity = num_unique_words / total_words
  ) %>%
  arrange(speech_complexity) %>% 
  slice_max(total_words, n = 4, with_ties = FALSE)

df_speech_complexity

df_speech_complexity %>% 
  mutate(raw_character_text = fct_reorder(raw_character_text, desc(speech_complexity))) %>% 
  ggplot(aes(raw_character_text, speech_complexity, fill = raw_character_text)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

#### Compute "star character" for each episode and their number of lines spoken

This script can be modified to find the top `x` characters per episode by changing the parameter `n` in `slice_max`.

Additionally, this code joins the total words per episode and computes the percentage of the script relative to each character

```{r}
get_word_counts_per_episode <- function(df_script) {
  df_script %>%
    filter(speaking_line == TRUE) %>%
    mutate(
      word_count = (as.numeric(word_count))
    ) %>%
    filter(!is.na(word_count)) %>%
    filter(word_count <= 1000) %>% 
    group_by(episode_id, raw_character_text) %>%
    summarize(total_word_count = sum(word_count), .groups = "drop") %>%
    group_by(episode_id) %>%
    slice_max(total_word_count, n = 1, with_ties = FALSE) %>%
    arrange(episode_id, desc(total_word_count))
}

episode_word_counts <- 
  get_word_counts_per_episode(df_script) %>% 
  inner_join(df_words_per_epsiode, by = "episode_id") %>% 
  select(c("episode_id", "raw_character_text", "total_word_count", "total_words_per_episode")) %>% 
  mutate(percentage_of_script = total_word_count / total_words_per_episode)
episode_word_counts

```

#### Join 'star character' data with the episode viewership data

```{r}
df_simpsons <- df_ep_new %>%
  left_join(episode_word_counts, by = "episode_id")

df_simpsons
```

#### First plot - select only the main characters and plot the star character in relation to viewership

```{r fig.width=9, fig.height=5}
# Bart|Homer|Marge|Lisa|Maggie|
df_simpsons %>%
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  ggplot(
    aes(x = `episode_id`, y = us_viewers_in_millions)
  ) +
  geom_point(aes(color = raw_character_text)) +
  geom_smooth(color = "black") +
  # geom_smooth(
  #   data = . %>%
  #     filter(raw_character_text == "Lisa Simpson"),
  #   color = "cornflowerblue"
  # ) +
  # geom_smooth(
  #   data = . %>%
  #     filter(raw_character_text == "Bart Simpson"),
  #   color = "lightcoral"
  # ) +
 labs(
   title = "Which Simpsons Characters Are Most Popular Over Time?",
   x = "Episode Number (sequential over time)",
   y = "US Viewership in Millions",
   color = '"Star Character" of Each Episode',
   caption = "The Star Character is conisdered the character with the most spoken words in a given episode."
 )
```

#### The above plot shows the overwhelming number of episodes where Homer is the star character. A trend line of Homer closely follows that of all the characters combined. How do the remaining 3 main characters perform?

```{r fig.width=9, fig.height=5}
# Bart|Homer|Marge|Lisa|Maggie|
df_simpsons %>%
  filter(grepl("Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  ggplot(
    aes(x = `episode_id`, y = us_viewers_in_millions)
  ) +
  geom_point(aes(color = raw_character_text)) +
  geom_smooth(color = "black", label="Overall Trend") +
  geom_smooth(aes(color = raw_character_text)) +
 labs(
   title = "Excluding Homer, Which Simpsons Characters Are Most Popular Over Time?",
   x = "Episode Number (sequential over time)",
   y = "US Viewership in Millions",
   color = '"Star Character" of Each Episode',
   caption = "The Star Character is conisdered the character with the most spoken words in a given episode."
 )
```

#### Create a similar graph in Boxplot form

```{r}
df_simpsons %>% 
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  #group_by(raw_character_text) %>% 
ggplot(
    aes(x = raw_character_text, y = us_viewers_in_millions, fill = raw_character_text)
  ) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

#### Compile data per year

```{r}
df_simpsons %>% 
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
ggplot(
    aes(x = `original_air_year`, y = us_viewers_in_millions, color = raw_character_text)
  ) +
  geom_count() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

#### Compute a IMDb votes per view - what episodes are people most passionate about?

Also add a smoothing line

```{r}
df_simpsons %>% 
  filter(grepl("Simpson|Santa's Little Helper|Obama|Trump|Millhouse|Chief Wiggum", raw_character_text)) %>% 
  mutate(votes_per_view = imdb_votes/us_viewers_in_millions) %>% 
ggplot(
    aes(x = `episode_id`, y = votes_per_view, color = raw_character_text)
    )+
  geom_point() +
  geom_smooth(color = "yellow")+
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```
