---
title: "Simpsons_Emery_cooking"
author: "Emery Lauer"
date: "2025-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# install.packages("plotly")
```

## EDA

```{r}
df_ep <- read.csv("data/simpsons_episodes.csv")
df_script <- read.csv("data/simpsons_script_lines.csv")

head(df_ep)
head(df_script)
```

#### Compute the total number of words per episode & plot relative to viewership

```{r}
# Clean up df_ep to make it slightly easier to join down the line (ie, have the join by colum be named the same thing)
# Also filter out extra columns that are not needed
df_ep_new <- df_ep %>%
  select(id, imdb_votes, original_air_date, original_air_year, us_viewers_in_millions, views, season) %>%
  rename(
    episode_id = id
  )

# Compute the total number of works in an episode
df_words_per_epsiode <-
  df_script %>%
  filter(speaking_line == TRUE) %>%
  mutate(word_count = as.numeric(word_count)) %>%
  filter(word_count <= 1000) %>%
  group_by(episode_id) %>%
  summarise(total_words_per_episode = sum(word_count)) %>%
  inner_join(df_ep_new, by = "episode_id")

df_words_per_epsiode %>%
  ggplot(aes(total_words_per_episode, us_viewers_in_millions)) +
  geom_point() +
  labs(
    x = "Number of Words in an Episode",
    y = "Episode Viewership (in millions of views)",
    title = "How does dialog quantity impact viewership of The Simpsons?",
    caption = "Source: The Simpsons Dataset, Prashant Banerjee (via Kaggle)"
  )
```

#### Compute words over time

```{r}
df_words_per_epsiode %>%
  ggplot(aes(episode_id, total_words_per_episode)) +
  geom_point() +
  geom_smooth() +
  labs(
    x = "Episode ID (sequential with time)",
    y = "Total Number of Words in Episode",
    title = "How has dialog quantity varied over time in The Simpsons?",
    caption = "Source: The Simpsons Dataset, Prashant Banerjee (via Kaggle)"
  )
```

#### Compute "star character" for each episode and their number of lines spoken

This script can be modified to find the top `x` characters per episode by changing the parameter `n` in `slice_max`.

Additionally, this code joins the total words per episode and computes the percentage of the script relative to each character

Finally, join this data with the episode data.

```{r}
get_word_counts_per_episode <- function(df_script) {
  df_script %>%
    filter(speaking_line == TRUE) %>%
    mutate(
      word_count = (as.numeric(word_count))
    ) %>%
    filter(!is.na(word_count)) %>%
    filter(word_count <= 1000) %>%
    group_by(episode_id, raw_character_text) %>%
    summarize(total_word_count = sum(word_count), .groups = "drop") %>%
    group_by(episode_id) %>%
    slice_max(total_word_count, n = 1, with_ties = FALSE) %>%
    arrange(episode_id, desc(total_word_count))
}

episode_word_counts <-
  get_word_counts_per_episode(df_script) %>%
  inner_join(df_words_per_epsiode, by = "episode_id") %>%
  select(
    c(
      "episode_id",
      "raw_character_text",
      "total_word_count",
      "total_words_per_episode"
    )
  ) %>%
  mutate(percentage_of_script = total_word_count / total_words_per_episode)

df_simpsons <- df_ep_new %>%
  left_join(episode_word_counts, by = "episode_id")

df_simpsons
```

#### First plot - select only the main characters and plot the star character in relation to viewership

```{r fig.width=9, fig.height=5}
# Bart|Homer|Marge|Lisa|Maggie|
df_simpsons %>%
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  ggplot(
    aes(x = `episode_id`, y = us_viewers_in_millions)
  ) +
  geom_point(aes(color = raw_character_text)) +
  geom_smooth(color = "black") +
  # geom_smooth(
  #   data = . %>%
  #     filter(raw_character_text == "Lisa Simpson"),
  #   color = "cornflowerblue"
  # ) +
  # geom_smooth(
  #   data = . %>%
  #     filter(raw_character_text == "Bart Simpson"),
  #   color = "lightcoral"
  # ) +
  labs(
    title = "Which Simpsons Characters Are Most Popular Over Time?",
    x = "Episode Number (sequential over time)",
    y = "US Viewership in Millions",
    color = '"Star Character" of Each Episode',
    caption = "The Star Character is conisdered the character with the most spoken words in a given episode. \n Source: The Simpsons Dataset, Prashant Banerjee (via Kaggle)"
  )
```

#### The above plot shows the overwhelming number of episodes where Homer is the star character. A trend line of Homer closely follows that of all the characters combined. How do the remaining 3 main characters perform?

```{r fig.width=9, fig.height=5}
# Bart|Homer|Marge|Lisa|Maggie|
df_simpsons %>%
  filter(grepl("Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  ggplot(
    aes(x = `episode_id`, y = us_viewers_in_millions)
  ) +
  geom_point(aes(color = raw_character_text)) +
  geom_smooth(color = "black", label = "Overall Trend") +
  geom_smooth(aes(color = raw_character_text)) +
  labs(
    title = "Excluding Homer, Which Simpsons Characters Are Most Popular Over Time?",
    x = "Episode Number (sequential over time)",
    y = "US Viewership in Millions",
    color = '"Star Character" of Each Episode',
    caption = "The Star Character is conisdered the character with the most spoken words in a given episode. \n Source: The Simpsons Dataset, Prashant Banerjee (via Kaggle)"
  )
```

#### Bart seems to be the second most popular in the earlier episodes but falls off the curve around episode 200. Lisa and Marge seem to perform better than the trend line in some cases. We are curious if there is any association with when female writers entered the field and the quality of these two main female characters.

```{r}
df_women_writers <- read.csv("data/women_writers.csv", header = FALSE)
colnames(df_women_writers) <- c("names", "seasons")

head(df_women_writers)

all_seasons <- unlist(strsplit(df_women_writers$seasons, ","))

all_seasons <- as.numeric(all_seasons)

season_counts <- as.data.frame(table(all_seasons))
colnames(season_counts) <- c("season", "num_female_writers")

season_counts$season <- as.numeric(as.character(season_counts$season))


print(season_counts)
```

```{r}
df_simpsons <- df_simpsons %>%
  select(-starts_with("num_female_writers")) %>%
  left_join(season_counts, by = "season")

season_ranges <- df_simpsons %>%
  filter(!is.na(num_female_writers)) %>%
  group_by(season, num_female_writers) %>%
  summarise(start_ep = min(episode_id), end_ep = max(episode_id), .groups = "drop")

ggplot(
  df_simpsons %>%
    filter(grepl("Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)),
  aes(x = episode_id, y = us_viewers_in_millions)
) +
  geom_rect(
    data = season_ranges,
    aes(
      xmin = start_ep, xmax = end_ep,
      ymin = -Inf, ymax = Inf,
      alpha = num_female_writers
    ),
    fill = "chartreuse3", inherit.aes = FALSE, show.legend = FALSE
  ) +
  geom_point(aes(color = raw_character_text)) +
  geom_smooth(color = "black", label = "Overall Trend") +
  geom_smooth(aes(color = raw_character_text)) +
  scale_alpha_continuous(range = c(0.05, 0.3)) +
  labs(
    title = "Excluding Homer, Which Simpsons Characters Are Most Popular Over Time?",
    x = "Episode Number (sequential over time)",
    y = "US Viewership in Millions",
    color = '"Star Character" of Each Episode',
    caption = "Red shaded areas = seasons with female writers; darker = more female writers.\nSource: The Simpsons Dataset, Prashant Banerjee (via Kaggle)"
  ) +
  theme_minimal()
```

#### Create a similar graph in boxplot form to better visualize how often each character has the most words in episodes

```{r}
df_simpsons %>%
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  # group_by(raw_character_text) %>%
  ggplot(
    aes(x = raw_character_text, y = us_viewers_in_millions, fill = raw_character_text)
  ) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

#### Compile data per year to see if world events affect viewership

```{r}
df_simpsons %>%
  filter(grepl("Homer Simpson|Bart Simpson|Marge Simpson|Lisa Simpson", raw_character_text)) %>%
  ggplot(
    aes(x = `original_air_year`, y = us_viewers_in_millions, color = raw_character_text)
  ) +
  geom_count() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

In the above figure, you may notice that there is a peak from 2000 to around 2003. This may seem like viewers picked the show back up after a decline, however it simply denotes a period when they changed how the viewership was counted.

#### Compute a IMDb votes per view - what episodes are people most passionate about?

Also add a smoothing line

```{r}
df_simpsons %>%
  filter(grepl("Simpson|Santa's Little Helper|Obama|Trump|Millhouse|Chief Wiggum", raw_character_text)) %>%
  mutate(votes_per_view = imdb_votes / us_viewers_in_millions) %>%
  ggplot(
    aes(x = `episode_id`, y = votes_per_view, color = raw_character_text)
  ) +
  geom_point() +
  geom_smooth(color = "yellow") +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))
```

#### Compute Speech Complexity

```{r fig.height=5}
unique_words <- function(line) {
  # Convert all characters to lowercase
  tolower(line) %>%
    # Remove all non-alphanumeric characters (ie punctuation)
    gsub("[^[:alnum:] ]", "", .) %>%
    # Split into individual words
    str_split(" +") %>%
    # Deal with weird list-in-a-list thing
    .[[1]] %>%
    # Count onlthe unique (non-repeat words)
    unique() %>%
    # Get total number of words
    length()
}

df_speech_complexity <-
  df_script %>%
  mutate(word_count = as.numeric(word_count)) %>%
  filter(word_count < 1000) %>%
  filter(
    grepl(
      "Simpson",
      raw_character_text
    )
  ) %>%
  group_by(raw_character_text) %>%
  summarise(
    all_words = paste0(normalized_text, collapse = " "),
    num_unique_words = unique_words(all_words),
    total_words = sum(word_count, na.rm = TRUE),
    speech_complexity = (num_unique_words / total_words) * 100
  ) %>%
  arrange(speech_complexity) %>%
  slice_max(total_words, n = 4, with_ties = FALSE)

df_speech_complexity

df_speech_complexity %>%
  mutate(raw_character_text = fct_reorder(raw_character_text, desc(speech_complexity))) %>%
  ggplot(aes(raw_character_text, speech_complexity, fill = raw_character_text)) +
  geom_col() +
  theme(
    axis.text.x = element_text(angle = 50, hjust = 1),
    plot.caption = element_text(hjust = 0)
  ) +
  labs(
    x = "Character Name",
    color = "Character Name",
    y = "Series-Wide Speech Complexity (percentage)",
    caption = "Speech Complexity is defined as the number of unique words as a percentage \n of the total words spoken in a character's dialog across the entire series. \n Source: The Simpsons Dataset, Prashant Banerjee (via Kaggle)",
    title = "Which Simpsons Main Characters Have the Most Complex Speech?"
  )
```

In the above graph, Lisa has the most complex work set while homer has the least. This is interesting because homer is often the character with the most words in most episodes and Lisa has the least.
